---
title: "Article 19 regional expenses, 2011–2015"
output:
  html_notebook: 
    code_folding: hide
    highlight: pygments
    theme: cosmo
---

Article 19 has generally increased its spending on regional programs evenly, with some notable exceptions. Prior to 2014, the bulk of its spending was directed to programs in Africa, wih a peak in 2013 of nearly £1,000,000. In the past couple years, however, African spending has been reduced, while global programs have increased, possibly as a replacement (though not entirely, since global programs were the second most funded category in 2013).

```{r, message=FALSE, echo=FALSE}
knitr::opts_chunk$set(fig.retina=2)
```


```{r, message=FALSE}
library(tidyverse)
library(ggstance)
library(forcats)

a19 <- read_csv("regional_expenses_2011-2015.csv",
                col_types=cols(
                  Year = col_integer(),
                  Subcategory = col_character(),
                  Category = col_character(),
                  Amount = col_integer()
                )) %>%
  filter(!(Category %in% c("Governance", "Fundraising", "Law and policy"))) %>%
  mutate(Category = ordered(fct_relevel(Category, "Global"))) %>%
  group_by(Year, Category) %>%
  summarise(Amount = sum(Amount))

a19.plot.amount <- ggplot(a19, aes(x=Year, y=Amount, colour=Category)) +
  geom_line(size=1) +
  guides(colour=guide_legend(title=NULL)) +
  labs(x=NULL, y=NULL,
       title="Article 19 annual expenditures on regional programs",
       subtitle="Amount spent (£)",
       caption="Source: 2011–2015 annual reports") +
  scale_y_continuous(labels=scales::dollar_format(prefix="£")) +
  theme_light(base_family="Source Sans Pro") +
  theme(legend.position="bottom")
a19.plot.amount
```

Looking at the percent of money spent on regional programs highlights this shift even better. Africa had been receiving roughly 30% of all funding until 2014, when Asia became more prominent. By 2015, all regions get between 12 and 22 percent.

```{r}
a19.percent <- a19 %>%
  group_by(Year) %>%
  mutate(total = sum(Amount),
         percent = Amount / total)

a19.plot.percent <- ggplot(a19.percent, aes(x=Year, y=percent, colour=Category)) +
  geom_line(size=1) +
  guides(colour=guide_legend(title=NULL)) +
  labs(x=NULL, y=NULL,
       title="Article 19 annual expenditures on regional programs",
       subtitle="% of total regional spending",
       caption="Source: 2011–2015 annual reports") +
  scale_y_continuous(labels=scales::percent) +
  theme_light(base_family="Source Sans Pro") +
  theme(legend.position="bottom")
a19.plot.percent
```

